---
title: "Discovery API Text Analysis"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: yes
    number_sections: no
    toc: yes
    toc_float: yes
---

```{r echo = FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, cache = TRUE,
               cache.lazy = FALSE, echo = TRUE)

options(width = 100, dplyr.width = 100, knitr.table.format = "html")
library(ggplot2)
library(methods)
ggplot2::theme_set(theme_light())
```


```{r load_packages, echo=FALSE, eval=FALSE}

packages <- c("kableExtra", "tidytext", "jsonlite", "tidyverse", "widyr", "igraph", "ggraph", "AzureML", "topicmodels")

#install.packages(packages)

lapply(packages,library, character.only = TRUE)

```

The Socrata data platform hosts tens of thousands of government datasets. Governments large and small publish data on crime, permits, finance, healthcare, research, performance, and more for citizens to use. While this large corpus of government data is already accessible via [opendatanetwork.com](opendatanetwork.com), this API is a powerful way to access and explore all public metadata published on the Socrata platform.

Using the [Socrata Discovery API](https://socratadiscovery.docs.apiary.io) I'll use metadata from these datasets to understand the connections between them. 

The metadata includes information like the name of the dataset, a description field, tags for the dataset that have been assigned by a human being, and so forth. The metadata for all its public datasets is [available online in JSON format](http://api.us.socrata.com/api/catalog/v1).

In this report, I will analyze the Socrata Open Data Network metadata as a text dataset and perform text mining techniques using the R library [tidytext](https://www.tidytextmining.com/). I will preform word co-occurrences and correlations, tf-idf, and topic modeling to explore the connections between the datasets. I will seek to find if datasets are related to one other and find clusters of similar datasets. Since the Socrata Open Data Network provides several text fields in the metadata, most importantly the name, description, and tag fields, I can show connections between the fields to better understand the connections between the Socrata Open Data Network API datasets.

## How data is organized at the Socrata Open Data Network

Download the JSON file and take a look at the names of what is stored in the metadata.

```{r download}
library(jsonlite)

metadata <- jsonlite::fromJSON("https://api.us.socrata.com/api/catalog/v1?limit=10000")

# look at column names in dataset
base::names(metadata$results)

```

The name, description, and  for each dataset will be the features of interest.

```{r class, dependson = "download"}

# text fields of interest are within the 'resource' and 'classification' object
base::class(metadata$results$resource$name)
base::class(metadata$results$resource$description)
base::class(metadata$results$classification$categories) # domain_tags are an array (or list)
```

The name and description fields are stored as character vectors, and the domain_tags are stored as a list of character vectors.

### Data preparation

```{r name, inlcude = FALSE, dependson = "download", message=FALSE}
## Create dataframes of title, description, and tags fields
library(tidyverse)

socrata_title <- dplyr::data_frame(
  id = metadata$results$resource$id,
  title = metadata$results$resource$name
)

# show title fields
socrata_title %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")
```



```{r desc, dependson = "download", dplyr.width = 150}
socrata_desc <- dplyr::data_frame(
  id = metadata$results$resource$id,
  desc = metadata$results$resource$description
  ) %>% 
  dplyr::filter(!purrr::map_lgl(desc, is.null)) %>% # filter out null values
  dplyr::filter(!purrr::map_lgl(desc, is.na)) %>% # filter out NA values
  dplyr::filter(desc != "") # filter out blank values
  
  
# take a sample and show description fields
socrata_desc %>%
  dplyr::select(desc) %>%
  dplyr::sample_n(5) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```


```{r domain_tag, dependson = "download"}
# all datasets have a title and description but not all have domain_tags
socrata_tags <- dplyr::data_frame(
  id = metadata$results$resource$id,
  tag = metadata$results$classification$domain_tags
  ) %>% 
  dplyr::filter(!purrr::map_lgl(tag, is.null)) %>% # filter out null values
  tidyr::unnest(tag) # unlist the tag array and make the dataframe long

# show tags fields
socrata_tags %>%
  dplyr::sample_n(100) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")
  
```


```{r unnest, dependson = c("title","desc")}
# Create dataframes of tokenized title, description, and tags fields
library(tidytext)

title_tokens <- socrata_title %>%
  tidytext::unnest_tokens(word, title) %>%
  dplyr::anti_join(stop_words, by = "word") # remove default stop words

desc_tokens <- socrata_desc %>%
  tidytext::unnest_tokens(word, desc) %>%
  dplyr::anti_join(stop_words, by = "word") # remove default stop words

# don't remove any stop words
# the thinking here is that these tags are selected by a user
# and have meaning to both the dataset and the user
# there is considerablly more thought placed on these tags than say,
# a title or description
# Put differently, these are not an arbitrary listing of tags 
# and thus, they should all remain
tag_tokens <- socrata_tags %>%
  tidytext::unnest_tokens(word, tag) # we're NOT removing any tags here

```

The name, description, and domain_tag datasets have been prepared and are now ready for exploration.

```{r extra_stopwords, echo = FALSE, dependson = "unnest"}

# create a list of user-defined stop words
# TO DO: finish adding extra stop words to this list
# TO DO: move this function into another R file
extra_stopwords <- dplyr::data_frame(
  word = c(
    as.character(0:310), # don't include 311
    as.character(312:910), # don't include 911
    as.character(912:9999),
    c("1st","2nd","3rd","4th","5th","6th","7th","8th","9th","10th","11th","12th","13th","14th","15th"),
    paste0("0", as.character(0:999)),
    paste0("00", as.character(0:9)),
    as.character(1950:2019), 
    "2013inspections","fhv","iii","insp","r.c.","since2012","ui","http","https","br","c7ck","chicagopolice.org",
    "obm.html","opendatatutorial.aspx","1kyje6x","csv","data.cityofchicago.org","data.cityofnewyork.us","href",
    "div","docs.google.com","en","1xrivpz","1cx6zvs","1dyglwj","2xku","438e","31st","296","1234","edit","assets",
    "206","236","296","309","360","2.3","4700","8844","428","877","428","2,900","312.745.6071","65,000","80,000",
    "1eeo5t_lt8qhghmjpj6m9071y5ofoxcadfi4f7aiz2ie","201503_cfpb_narrative","files.consumerfinance.gov","sets",
    "h.rjwrosiwiw5o","i.e.","ingress.kingcounty.gov","standard.pdf","tlc.nyc.gov","webapps.cityofchicago.org",
    "www.cityofchicago.org","www.consumerfinance.gov","www.health.ny.gov","www.ilhousingsearch.org","dataset",
    "www.kingcounty.gov","www.nyc.gov","yqxq","on.ny.gov","searchaddresspage.html","work.shtml","i.e","data",
    "document","documentation","schema","heading","created","record","set","inspection_serial_num","row","rows",
    "files","complaintdatabase","vital_records","genealogy","nbsp","amp","target","_blank","5yearweeklyaverage.pdf",
    "_appendix_reduced_for_web.pdf","acs_field_description_dictionary.xlsx","adult_indicators.html","agdistricts.html",
    "annualleafcollection.html","attachment1_e.html","authenticationchooser.html","bacppublicvehicles.html",
    "benchmarking_id_request.html","building_violationsonline.html","changes.html","chsxindex.html","cou.html",
    "children_s_safetyzoneporgramautomaticspeedenforcement.html","description.html","downloads.html",
    "contact_us_e.html","data_disclaimer.html","lobby.html","portal.html","light_cameraenforcement.html","item.html",
    "intro.html","index.html","educationalresources.html","fluoridation_indicators.html","prohibited_buildings_list_affidavit.html",
    "public_chauffeurinformation.html","state.html","streets.html","apps.health.ny.gov","apps.nccd.cdc.gov","austintexas.gov",
    "cdc.gov","census.gov","cde.ca.gov","chattanooga.gov","checkbook.iowa.gov","chhs.data.ca.gov","chronicdata.cdc.gov",
    "comptroller.texas.gov","2fmaps.nccs.nasa.gov","access.nyc.gov","accesshub.pdc.wa.gov","apps.suffolkcountyny.gov",
    "blightstatus.nola.gov","broadbandmap.fcc.gov","budget.kcmo.gov","cdph.data.ca.gov","cityofboston.gov",
    "cmadmin.energystar.gov","cookcountyil.gov","council.nyc.gov","cpa.texas.gov","data.austintexas.gov",
    "data.cambridgema.gov","data.cdc.gov","data.cityofboston.gov","data.cms.gov","data.ct.gov","data.detroitmi.gov",
    "data.energystar.gov","data.govloop.com","data.healthcare.gov","data.kiingcounty.gov","data.maryland.gov",
    "data.melbourne.vic.gov.au","data.mo.gov","data.montgomerycountymd.gov","data.nasa.gov","data.nola.gov",
    "data.ny.gov","data.oregon.gov","data.pr.gov","data.raleighnc.gov","data.seattle.gov","data.texas.gov",
    "data.transportation.gov","data.vermont.gov","data.wa.gov","datacatalog.cookcountyil.gov","detroitmi.gov",
    "developer.nrel.gov","directory.osd.gss.omb.delaware.gov","dmh.mo.gov","edsight.ct.gov",
    "electionsdata.kingcounty.gov","energy.maryland.gov","ephtracking.cdc.gov","fcc.gov","firstmap.delaware.gov",
    "fortress.wa.gov","ftp.cdc.gov","ftp.census.gov","geoportal.igac.gov.co","gis.oregon.gov",
    "gisrevprxy.seattle.gov","governor.ny.gov","hartford.gov","hawaii.gov","health.data.ny.gov","healthcare.gov",
    "healthpsi.nyc.gov","hub.nashville.gov","innovation.cms.gov","innovations.cms.gov","its.ny.gov",
    "justice.oregon.gov","kcmo.gov","kids.delaware.gov","kingcounty.gov","lacounty.gov","malegislature.gov",
    "maps.nyc.gov","maps.raleighnc.gov","maryland.gov","mcit.gov.co","medicaid.gov","melbourne.vic.gov.au",
    "miamidade.gov","michigan.gov","montgomerycountymd.gov","msc.fema.gov","namus.gov","nashville.gov",
    "nationalservice.gov","nc.gov","neo.jpl.nasa.gov","nola.gov","nyc.gov","oce.pr.gov","ocfs.ny.gov","oh.gov",
    "onlinedevcenter.raleighnc.gov","open.ny.gov","opencheckbook.ct.gov","opencheckbook.providenceri.gov",
    "opendata.lasvegasnevada.gov","openpaymentsdata.cms.gov","oregon.gov","orthos.dhses.ny.gov","pcip.gov",
    "permits.performance.gov","policia.pr.gov","portal.ct.gov","ppair.uspto.gov","profiles.health.ny.gov",
    "results.nola.gov","seattle.gov","sonomacounty.ca.gov","spending.dbm.maryland.gov","tools.cdc.gov",
    "ucr.fbi.gov","usa.gov","water.epa.gov","wonder.cdc.gov","www.abs.gov.au","www.afdc.energy.gov",
    "www.agriculture.ny.gov","www.austintexas.gov","www.boyaca.gov.co","www.broadbandmap.gov",
    "www.broadbandmap.ny.gov","www.budget.delaware.gov","www.cambridgema.gov","www.cdc.gov","www.census.gov",
    "www.cms.gov","www.colorado.gov","www.cookcountyil.gov","www.data.act.gov.au","www.datos.gov.co",
    "www.dec.ny.gov","www.detroitmi.gov","www.dhcs.ca.gov","www.dmhc.ca.gov","www.dnr.mo.gov",
    "www.doc.delaware.gov","www.dot.ny.gov","www.dwa.gov.za","www.ecy.wa.gov","www.elections.ny.gov",
    "www.energystar.gov","www.fbi.gov","www.fcc.gov","www.fns.usda.gov","www.ftb.ca.gov","www.governor.ny.gov",
    "www.govloop.com","www.hud.gov","www.illinoishistory.gov","www.imls.gov","www.labormarketinfo.ca.gov",
    "www.mainroads.wa.gov.au","www.mdot.maryland.gov","www.medicaid.gov","www.melbourne.vic.gov.au",
    "www.montgomerycountymd.gov","www.nola.gov","www.ny.gov","www.nyhealth.gov","www.nyserda.ny.gov",
    "www.open.ny.gov","www.oregon.gov","www.oria.wa.gov","www.permits.performance.gov","www.raleighnc.gov",
    "www.redmond.gov","www.revenue.pa.gov","www.seattle.gov","www.spending.dbm.maryland.gov","www.stpaul.gov",
    "www.tax.ny.gov","www.tdlr.texas.gov","www.transport.act.gov.au","www.transportation.gov","www.uspto.gov",
    "wwwn.cdc.gov","your.kingcounty.gov","yourmoney.nj.gov","_dam_may_2016_meta_data_info.pdf",
    "_document_reduced_for_web.pdf","2014graffitivandalismauditreport.pdf","20description.pdf",
    "20documentation.pdf","20documentation.pdf","appendices.pdf","area.pdf","bmmap.pdf",
    "caseload_dec_2016_current_zip.pdf","chsmethods.pdf","cityracks_map_metadata.pdf",
    "data_dictionary_trip_records_green.pdf","data_dictionary_trip_records_yellow.pdf",
    "dataset_description_06_10_portal_only.pdf","dataset_description_languages_2012_for_portal_only.pdf",
    "dataset_description_socioeconomic_indicators_2012_for_portal_only.pdf","dbs_brfss_survey.pdf",
    "definitions.pdf","kbyf_brochure.pdfthis","leadtestingregs.pdf","measures.pdf",
    "mobile_photo_enforcement_faq.pdf","nat1999doc.pdf","newdriverappstatuslookuplegend.pdf","nfirs_codes.pdf",
    "opendatahandbook.pdf","permittimelinessprogressreport.pdf","plan.pdf","prescriber_methods.pdf","readme.pdf",
    "selfac_datainfo.pdf","sex_race_ethnicity.pdf","survey.pdf","usvss.pdf","wide_flood_mitigation_study.pdf",
    "data_quality_e.html","demolition_delay.html","ethics.html","find_a_clinic.html","fshome.html","fss.html",
    "ghg.html","quality_measures_standards.html","rates_deaths_byage.html","riseordinance.html","securedtaxes.html",
    "services.html","sidewalk_cafe_current_permits.html","ucd.html","viewer.html","wages.html","yrbsindex.html",
    "_door","_historical_reports","_info_","_neat","_share","_shelter","_take"
  )
)
```

```{r reduced_tokens, dependson = "extra_stopwords"}

# remove those extra stop words from title and description
title_tokens_reduced <- title_tokens %>%
  dplyr::anti_join(extra_stopwords, by = "word")

desc_tokens_reduced <- desc_tokens %>%
  dplyr::anti_join(extra_stopwords, by = "word")

tag_tokens_reduced <- tag_tokens %>%
  dplyr::anti_join(dplyr::data_frame(word = c("and","of")), by = "word") # just remove 'and' and 'of'

title_tokens_reduced %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

desc_tokens_reduced %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

tag_tokens_reduced %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```


### Initial simple exploration

What are the most common words in the Socrata Open Data Network dataset domain_tags?

```{r dependson = "unnest"}
#What are the most common tags?
tag_tokens_reduced %>%
  dplyr::group_by(word) %>%
  dplyr::count(sort = TRUE) %>%
  dplyr::filter(n > 250) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

What are the most common words in the Socrata Open Data Network dataset descriptions?

```{r dependson = "reduced_tokens"}
#What are the most common descriptive words?
desc_tokens_reduced %>%
  dplyr::group_by(word) %>%
  dplyr::count(sort = TRUE) %>%
  dplyr::filter(n > 750) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

What are the most common words in the Socrata Open Data Network dataset titles?

```{r dependson = "reduced_tokens"}
#What are the most common words in titles?
title_tokens_reduced %>%
  dplyr::group_by(word) %>%
  dplyr::count(sort = TRUE) %>%
  dplyr::filter(n > 150) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## Word co-ocurrences and correlations

Here I examine which words commonly occur together in the titles, descriptions, and domain_tags of the Socrata Open Data Network datasets to create word networks that help determine which datasets are related to one other. 

### Networks of Description and Title Words

```{r title_word_pairs, dependson = "reduced_tokens"}
library(widyr)

title_word_pairs <- title_tokens_reduced %>%
  widyr::pairwise_count(word, id, sort = TRUE, upper = FALSE)

title_word_pairs %>%
  dplyr::arrange(-n) %>%
  dplyr::top_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```

These are the pairs of words that occur together most often in title fields. 

```{r desc_word_pairs, dependson = "reduced_tokens"}
desc_word_pairs <- desc_tokens_reduced %>%
  widyr::pairwise_count(word, id, sort = TRUE, upper = FALSE)

desc_word_pairs %>%
  dplyr::arrange(-n) %>%
  dplyr::top_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```

These are the pairs of words that occur together most often in description fields. 

Below is a plot of networks of these co-occurring words to better see relationships.

```{r plottitle, dependson = "title_word_pairs", fig.height=6, fig.width=9, fig.cap="Word network in the Socrata Open Data Network dataset titles"}
library(ggplot2)
library(igraph)
library(ggraph)

# plot network of co-occuring words for 'title' field
set.seed(1234)
title_word_pairs %>%
  dplyr::filter(n >= 50) %>%
  igraph::graph_from_data_frame() %>%
  ggraph::ggraph(layout = "fr") +
  ggraph::geom_edge_link(
    ggplot2::aes(edge_alpha = n, edge_width = n),
    edge_colour = "steelblue"
  ) +
  ggraph::geom_node_point(size = 5) +
  ggraph::geom_node_text(
    ggplot2::aes(label = name), 
    repel = TRUE, 
    point.padding = unit(0.2, "lines")
  ) +
  ggplot2::theme_void()
```

We see some clear clustering in this network of title words; words in the Socrata Open Data Network dataset titles are largely organized into several families of words that tend to go together.

Now I plot the same for the description fields.

```{r plotdesc, dependson = "desc_word_pairs", fig.height=6, fig.width=9, fig.cap="Word network in the Socrata Open Data Network dataset descriptions"}

# plot network of co-occuring words for 'description' field
set.seed(1234)
desc_word_pairs %>%
  dplyr::filter(n >= 260) %>%
  igraph::graph_from_data_frame() %>%
  ggraph::ggraph(layout = "fr") +
  ggraph::geom_edge_link(
    ggplot2::aes(edge_alpha = n, edge_width = n),
    edge_colour = "steelblue"
  ) +
  ggraph::geom_node_point(size = 5) +
  ggraph::geom_node_text(
    ggplot2::aes(label = name), 
    repel = TRUE, 
    point.padding = unit(0.2, "lines")
  ) +
  ggplot2::theme_void()
```

There seem to be two large network clusters of words around two phrases: "information" and  "world bank". 

### Networks of Tags

```{r tag_word_pairs, dependson = "reduced_tokens"}
tag_word_pairs <- tag_tokens_reduced %>%
  widyr::pairwise_count(word, id, sort = TRUE, upper = FALSE)

tag_word_pairs %>%
  dplyr::arrange(-n) %>%
  dplyr::top_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```

These are the pairs of words that occur together most often in domain_tags fields. 

```{r tag_wor_corr, dependson = "reduced_tokens"}

# find the correlations among tags
tag_word_corr <- tag_tokens_reduced %>%
  dplyr::group_by(word) %>%
  dplyr::filter(n() >= 100) %>%
  widyr::pairwise_cor(word, id, sort = TRUE, upper = FALSE)

tag_word_corr %>%
  dplyr::arrange(-correlation) %>%
  dplyr::top_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")


```

These are the most correlated pairs of words in domain_tags fields. 

Now I plot the same for the domain_tags fields.

```{r plottags, dependson = "tag_word_pairs", fig.height=6, fig.width=9, fig.cap="Word network in the Socrata Open Data Network dataset domain tags"}

# plot network of co-occuring words for 'doamin_tags' field
set.seed(1234)
tag_word_pairs %>%
  dplyr::filter(n >= 100) %>%
  igraph::graph_from_data_frame() %>%
  ggraph::ggraph(layout = "fr") +
  ggraph::geom_edge_link(
    ggplot2::aes(edge_alpha = n, edge_width = n),
    edge_colour = "steelblue"
  ) +
  ggraph::geom_node_point(size = 5) +
  ggraph::geom_node_text(
    ggplot2::aes(label = name), 
    repel = TRUE, 
    point.padding = unit(0.2, "lines")
  ) +
  ggplot2::theme_void()
```

Here we have another sparse, tightly-grouped network of terms that seem to logically relate to one another. For instance, "elections" and "results" have a strong connection, which leads us to conclude that many of the datasets contain only election results. Whereas "country", "beneficiary", and "level" are three significant nodes within a looser network relationship, suggesting that World Bank and CMS datasets -- while never overlapping -- otherwise overwhelm the corpus of datasets on the Discovery API.

As an exercise, I plot the network of correlated words as well.

```{r plotcorr, dependson = "tag_word_corr", fig.height=6, fig.width=9, fig.cap="Word network in the Socrata Open Data Network dataset domain tags"}

# plot network of correlated words for 'doamin_tags' field
set.seed(1234)
tag_word_corr %>%
  dplyr::filter(correlation > 0.6) %>%
  igraph::graph_from_data_frame() %>%
  ggraph::ggraph(layout = "fr") +
  ggraph::geom_edge_link(
    ggplot2::aes(edge_alpha = correlation, edge_width = correlation),
    edge_colour = "steelblue"
  ) +
  ggraph::geom_node_point(size = 5) +
  ggraph::geom_node_text(
    ggplot2::aes(label = name), 
    repel = TRUE, 
    point.padding = unit(0.2, "lines")
  ) +
  ggplot2::theme_void()
```

## Calculating tf-idf for the description fields

### Getting the tf-idf for the description field words

What are the highest tf-idf words in the Socrata Open Data Network description fields?

```{r desc_tf_idf, dependson = "reduced_tokens"}
library(topicmodels)

desc_tf_idf <- desc_tokens_reduced %>% 
  dplyr::count(id, word, sort = TRUE) %>%
  dplyr::ungroup() %>%
  tidytext::bind_tf_idf(word, id, n) %>%
  dplyr::arrange(-tf_idf)

desc_tf_idf %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```

These are the most important words in the description fields as measured by tf-idf, meaning they are common but not too common. 

### Connecting description fields to keywords

First join the results of the tf-idf analysis with the tags dataset.

```{r full_join, dependson = c("desc_tf_idf", "keyword")}
library(topicmodels)

# join tags on description tf-idf
desc_tf_idf_tags <- dplyr::full_join(
  desc_tf_idf, 
  socrata_tags , by = "id") %>%
  dplyr::arrange(word)

desc_tf_idf_tags %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```

Plot some of the most important words, as measured by tf-idf, for all of the provided keywords used on the Socrata Open Data Network datasets.

```{r plottfidf, dependson = "full_join", fig.width=10, fig.height=7, fig.cap="Distribution of tf-idf for words from datasets labeled with selected keywords"}

desc_tf_idf_tags %>% 
  dplyr::filter(!near(tf, 1)) %>%
  dplyr::filter(tag %in% c("health","safety","food","fire","rescue","closures")) %>%
  dplyr::arrange(dplyr::desc(tf_idf)) %>%
  dplyr::group_by(tag) %>%
  dplyr::distinct(word, tag, .keep_all = TRUE) %>%
  dplyr::top_n(15, tf_idf) %>% 
  dplyr::ungroup() %>%
  dplyr::mutate(word = base::factor(word, levels = base::rev(unique(word)))) %>%
  ggplot2::ggplot(ggplot2::aes(word, tf_idf, fill = tag)) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::facet_wrap(~tag, ncol = 3, scales = "free") +
  ggplot2::coord_flip() +
  ggplot2::labs(title = "Highest tf-idf words in Socrata metadata description fields",
       caption = "Socrata metadata from https://api.us.socrata.com/api/catalog/v1",
       x = NULL, y = "tf-idf")

```

## Uncovering hidden conversations
Topic modeling attempts to uncover the hidden conversations within each description field. [Latent Dirichlet allocation (LDA)](https://en.wikipedia.org/wiki/Latent_Dirichlet_allocation) is a technique to model each document (description field) as a mixture of topics and further describe each topic as a mixture of words

``` {r desc_word_counts}
desc_word_counts <- desc_tokens_reduced %>%
  dplyr::count(id, word, sort = TRUE) %>%
  dplyr::ungroup()

desc_word_counts %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```

### Casting to a document-term matrix

Create a sparse document term matrix, containing the count of terms in each document.

``` {r sparse_matrix}
desc_dtm <- desc_word_counts %>%
  tidytext::cast_dtm(id, word, n)

desc_dtm
```

### Ready for topic modeling

The following creates an LDA model. Like many clustering algorithms, the number of topics must be set *a priori*. Here I set the number of topics to 8.

```{r desc_lda, dependson = "sparse_matrix"}
library(topicmodels)
# run an LDA on the description words
desc_lda <- topicmodels::LDA(desc_dtm, k = 8, control = base::list(seed = 1234))
desc_lda
```

### Interpreting the topic model

The following takes the lda model and constructs a tidy data frame that summarizes the results.

```{r tidy_lda, dependson = "desc_lda"}
# interpret the results
tidy_lda <- tidytext::tidy(desc_lda)

tidy_lda %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```

The column $\beta$ shows the probability of that term being generated from that topic for that document. It is the probability of that term (word) belonging to that topic. 

The following examines the top 6 terms for each topic.

```{r top_lda_tags, dependson = "tidy_lda"}
top_lda_tags <- tidy_lda %>%
  dplyr::group_by(topic) %>%
  dplyr::top_n(10, beta) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(topic, -beta)

top_lda_tags %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```

Here are the results of the `top_lda_tags` exercise depicted visually:

```{r plotbeta, dependson = "top_lda_tags", fig.width=12, fig.height=16, fig.cap="Top terms in topic modeling of Discovery API metadata description field texts"}

top_lda_tags %>%
  dplyr::mutate(term =  stats::reorder(term, beta)) %>%
  dplyr::group_by(topic, term) %>%    
  dplyr::arrange(dplyr::desc(beta)) %>%  
  dplyr::ungroup() %>%
  dplyr::mutate(term = base::factor(base::paste(term, topic, sep = "__"), 
                       levels = base::rev(base::paste(term, topic, sep = "__")))) %>%
  ggplot2::ggplot(ggplot2::aes(term, beta, fill = base::as.factor(topic))) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::coord_flip() +
  ggplot2::scale_x_discrete(labels = function(x) base::gsub("__.+$", "", x)) +
  ggplot2::labs(
    title = "Top 10 terms in each LDA topic",
    x = NULL, y = base::expression(beta)) +
  ggplot2::facet_wrap(~ topic, ncol = 4, scales = "free")

```

The most frequently occuring terms in each of the topics tend to surround the concepts of food inspections, health quality indicators, World Bank datasets, and building permits.

The following examines topics that are associated with which description fields (i.e., documents). The probability, $\gamma$, is the probability that each document belongs in each topic.

```{r lda_gamma, dependson = "desc_lda"}
# examine which topics are associated with which description fields
lda_gamma <- tidytext::tidy(desc_lda, matrix = "gamma")

lda_gamma %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```

The variable, $\gamma$, has values that run from 0 to 1. Documents with values near zero means that those documents do not belong in each topic. Values with $\gamma$ values close to 1 indicate that these documents do belong in those topics.

This distribution below shows that most documents either do belong or do not belong to a given topic.

```{r plotgammaall, dependson = "lda_gamma", fig.width=7, fig.height=5, fig.cap="Probability distribution in topic modeling of Discovery API metadata description field texts"}
ggplot2::ggplot(lda_gamma, ggplot2::aes(gamma)) +
  ggplot2::geom_histogram(bins = 48) +
  ggplot2::scale_y_log10() +
  ggplot2::labs(title = "Distribution of probabilities for all topics",
       y = "Number of documents", x = base::expression(gamma))

```

The following plot shows how the probabilities are distributed within each topic:

```{r plotgamma, dependson = "lda_gamma", fig.width=10, fig.height=12, fig.cap="Probability distribution for each topic in topic modeling of Discovery API metadata description field texts"}
ggplot2::ggplot(lda_gamma, ggplot2::aes(gamma, fill = base::as.factor(topic))) +
  ggplot2::geom_histogram(bins = 16, show.legend = FALSE) +
  ggplot2::facet_wrap(~ topic, ncol = 4) +
  ggplot2::scale_y_log10() +
  ggplot2::labs(title = "Distribution of probability for each topic",
       y = "Number of documents", x = base::expression(gamma))
```

### Connecting topic modeling with keywords

The following connects topic models with the domain tags to see what relationships exist between the human-tagged terms and the modeled topics.

```{r lda_join, dependson = "lda_gamma"}

# join tags on description LDA
desc_lda_tags <- dplyr::full_join(
  lda_gamma, 
  socrata_tags, by = c("document" = "id"))

desc_lda_tags %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```


I keep the document-topic entries that have probabilities ($\gamma$) greater than 0.9.

```{r top_lda_gamma_tags, dependson = "lda_join"}

top_lda_gamma_tags <- desc_lda_tags %>% 
  dplyr::filter(!purrr::map_lgl(tag, is.null)) %>% # filter out null values
  dplyr::filter(!purrr::map_lgl(tag, is.na)) %>% # filter out NA values
  dplyr::filter(tag != "") %>% # filter out blank values
  dplyr::filter(gamma > 0.9) %>% 
  dplyr::count(topic, tag, sort = TRUE)

top_lda_gamma_tags %>%
  dplyr::sample_n(10) %>%
  knitr::kable() %>%
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  kableExtra::scroll_box(height = "300px")

```


The following plot shows the top tags for each topic.

```{r plottopkeywords, dependson = "top_keywords", fig.width=16, fig.height=16, fig.cap="Top tags in topic modeling of Discovery API metadata description field texts"}

top_lda_gamma_tags %>%
  dplyr::group_by(topic) %>%
  dplyr::top_n(5, n) %>%
  dplyr::group_by(topic, tag) %>%
  dplyr::arrange(desc(n)) %>%  
  dplyr::ungroup() %>%
  dplyr::mutate(
    tag = base::factor(
      base::paste(
        tag, topic, sep = "__"
      ),
      levels = base::rev(
        base::paste(
          tag, topic, sep = "__"
        )
      )
    )
  ) %>%
  ggplot2::ggplot(ggplot2::aes(tag, n, fill = base::as.factor(topic))) +
  ggplot2::geom_col(show.legend = FALSE) +
  ggplot2::labs(
    title = "Top tags for each LDA topic",
    x = NULL, y = "Number of documents"
  )+
  ggplot2::coord_flip() +
  ggplot2::scale_x_discrete(labels = function(x) gsub("__.+$", "", x)) +
  ggplot2::facet_wrap(~ topic, ncol = 4, scales = "free")
```

Discovery API datasets are tagged with keywords by human beings and the LDA topic model used in this analysis answers the question, "For the datasets with description fields that have a high probability of belonging to a given topic, what are the most common human-assigned keywords?"

## Summary

The Socraata Discovery API houses a diverse set of tables that shows, through a combination of network analysis, tf-idf, and topic modeling, that many are connected to one other despite the many institutions publishing datasets.
